### 2019-02-16 manuscript/methods.tex
- %B done with this iteration of methods.tex

If ploidy and breeding system each influence lineage diversification individually, it is logical to examine their possible joint effects. %B: unsure about edit, which was: "If both ploidy and breeding system have the potential to influence lineage diversification, we should logically consider their effects jointly."
We thus fit a multi-state model that includes both traits \citep[MuSSE,][]{fitzjohn_2012}.
The three states in this model are self-incompatible diploids (ID), self-compatible diploids (CD), and polyploids, which are always self-compatible (CP).
As explained above, we did not include a state for self-incompatible polyploids because they are not observed in the data, and   that trait combination state is mechanistically predicted not to occur.
We call this the \textit{ID/CD/CP ploidy and breeding system} model.
The model has ten parameters, six for diversification in each state ($\lambda_{ID},\ \lambda_{CD},\ \lambda_{CP}$ for speciation, $\mu_{ID},\ \mu_{CD},\ \mu_{CP}$ for extinction) and four for transitions between states ($\rho_I,\ \rho_C$ for polyploidization transitions from $ID$ and $CD$ to $CP$, respectively; $\delta$ for diploidization from $CP$ to $CD$; $q_{IC}$ for loss of self-incompatibility without polyploidization, from $ID$ to $CD$).
The total rate of loss of self-incompatibility, \ie transitions out of $ID$, is $q_{IC} + \rho_I$.
Diploidization from $CP$ to $ID$ is not allowed because it would represent a simultaneous regain of SI.

@@

For all four models that consider ploidy changes, we allowed that diploidization could happen.
Previous modeling approaches \citep{mayrose_2011} have argued against inferring diploidization rates when using ploidy data that comes from classifications based on chromosome number multiplicity or chromosome number change models like chromEvol \citep{mayrose_2010}.
These types of classifications do not allow for a ploidy reversion.
Where indicated, the classification of ploidy for the data used in our models was based on chromosome multiplicity at the genus level.
% ref data/sources table?
However, the majority of the ploidy classifications were adopted from original studies with alternative sources of information (\eg geographic distribution, genus ploidy distribution) where ploidy was defined by authors that found evidence for it.
Since it is not clear whether diploidization can be detected under alternative ploidy classifications or even classifications based on chromosome number multiplicity at the genus level, we also fit the models without diploidization in order to test  whether the conclusions about diversification are sensitive to including diploidization.
As discussed by \citet{servedio_2014}, the presence or absence of a hypothesis can have an exploratory goal.
In our case the diploidization parameter (or its absence, $\delta=0$) in our models is an opportunity to explore an assumption that might be important but that is not the single definitive process to understand the interactions among polyploidy, breeding system, and diversification.

@@

Parameters for each of the ten diversification models were performed using custom code in the RevBayes \citep{hoehna_2016} environment.
Code and results for all analyses is available at \url{https://github.com/roszenil/solploidy}.
For all analyses, a correction for incomplete sampling was included, based on assuming that the Solanaceae family has approximately 3,000 species ($s=595/3000$) as estimated by the Solanaceae Source project \citep{solsource}.
For all ten models, we assumed that speciation and extinction parameters  had log-normal prior distributions with means equal to the expected net diversification rate $(\text{number of taxa} / [2 \times \text{root age}])$ and standard deviation $0.5$.
Priors for parameters defining trait changes were assumed to be gamma distributed with parameters $k=0.5$ and $\theta=1$. 
For each model, an MCMC chain was run for 96 hours in the high-performance computational cluster at Minnesota Supercomputing Institute, which allowed for 5,000 generation of burn-in and a minimum of 200,000 generations of MCMC for each of the 6 models. 
For each model, convergence and mixing of the MCMC was tested using the R package coda and the software Tracer (see supplementary information for convergence plots). % eventually, add supp info fig ref

@@

We calculated the marginal likelihood for each of the ten models in RevBayes \citep{hoehna_2016}.
Marginal likelihoods were calculated using 50 stepping stone steps under the methodology of \citet{xie_2010}.
Each stepping stone step was found by calculating 500 generations of burn-in followed by a total of 1000 MCMC steps (\cref{table:marginallike}).
The calculation of each marginal likelihood ran for 24 hours on a high-performance computational cluster.

Using the marginal likelihood values, we calculated thirteen different Bayes factors. % E: it said "fourteen" but I only count thirteen?
Six compared the models of polyploidy against each other (D/P and D/P--A/B, each with or without diploidization), one compared the breeding system models (I/C and I/C--A/B), and six compared the models with both traits (ID/CD/CP and ID/CD/CP--A/B, each with or without diploidization) (\cref{table:bayesfactors}).
%
Other comparisons between these models are not valid because the input data are different under the different state space codings (\cref{figure:stateclassifications}).
In essence, the D/P, I/C, and ID/CD/CP state spaces are not lumpable with respect to one another \citep{tarasov_2019}.
%R: Emma this is where the figure can be important to show that we have to make different input
% E: I think this short explanation is clear enough, especially with the figure.  I'm not sure whether the hierarchical concept is exactly applicable here.

@@

\begin{table}
\begin{tabular}{@{}llcccccc@{}} \toprule
%\multicolumn{4}{r}{Models} \\ \cmidrule(r){3-5}
Model& Model& Ploidy & Diploidization & Breeding  & Hidden & Parameters & Marginal \\
& Type & &  &System & State & & Log- Likelihood \\
1. D/P &BiSSE &	Yes  &	Yes &	No	&No	& 6	& -1182.93 \\
2. D/P no $\delta$ & BiSSE  &	Yes & 	No	& No	& No & 	5 &	-1193.66\\
3. D/P- A/B  & HiSSE &	Yes &	Yes	&No &	Yes &	11 &	\textbf{-1145.69}\\
4. D/P no $\delta$-A/B &	HiSSE &	Yes & No &	No &	Yes &	10	&-1150.99\\
5. I/C &	BiSSE &	No &	 No	&Yes &	No &	 5 &  -1194.80 \\
6. I/C-A/B &	HiSSE &	No &	 No	&Yes &	Yes	& 10 & \textbf{-1155.37}\\
7. ID/P/CD & MuSSE &	Yes & 	Yes &	Yes &	No &	10 & -1344.50\\
8. ID/P/CD no $\delta$ &	MuSSE &	Yes & 	No &	Yes	&No &	9 &-1345.87\\
9. ID/P/CD-A/B & MuHiSSE &	Yes 	&Yes &	Yes &	Yes &	16 & \textbf{-1300.35} \\
10.ID/P/CD no $\delta$-A/B &	 MuHiSSE	 & Yes & 	No	&Yes &	Yes	&15 &-1303.55 \\ \bottomrule
\end{tabular}
\caption{Marginal likelihoods for the ten diversification models proposed.}
\label{table:marginallike}
\end{table}

>>>>

If ploidy and breeding system each influence lineage diversification individually, it is logical to examine their possible joint effects. %B: unsure about edit, which was: "If both ploidy and breeding system have the potential to influence lineage diversification, we should logically consider their effects jointly."
We thus fit a multi-state model that includes both traits \citep[MuSSE,][]{fitzjohn_2012}.
The three states in this model are self-incompatible diploids (ID), self-compatible diploids (CD), and polyploids, which are always self-compatible (CP).
As explained above, we did not include a state for self-incompatible polyploids because they are not observed in the data, and   that trait combination state is mechanistically predicted not to occur.
We call this the \textit{ID/CD/CP ploidy and breeding system} model.
The model has 10 parameters, six for diversification in each state ($\lambda_{ID},\ \lambda_{CD},\ \lambda_{CP}$ for speciation, $\mu_{ID},\ \mu_{CD},\ \mu_{CP}$ for extinction) and four for transitions between states ($\rho_I,\ \rho_C$ for polyploidization transitions from $ID$ and $CD$ to $CP$, respectively; $\delta$ for diploidization from $CP$ to $CD$; $q_{IC}$ for loss of self-incompatibility without polyploidization, from $ID$ to $CD$).
The total rate of loss of self-incompatibility, \ie transitions out of $ID$, is $q_{IC} + \rho_I$.
Diploidization from $CP$ to $ID$ is not allowed because it would represent a simultaneous regain of SI.

@@

%B: this section needs a motivation. Right now, that appears to be hung on an argument from a citation (mayrose_2010), which is not clarified. I think you guys may have the paleo-WGD literature in mind here. If so, perhaps simply stating something ought to be written about the expectations for capturing some underlying process we think is ongoing.
For all four models that consider ploidy changes, we allowed diploidization.
Previous modeling approaches \citep{mayrose_2011} have argued against inferring diploidization rates when using ploidy data that comes from classifications based on chromosome number multiplicity or chromosome number change models like chromEvol \citep{mayrose_2010}.
These types of classifications do not allow for a ploidy reversion.
Where indicated, the classification of ploidy for the data used in our models was based on chromosome multiplicity at the genus level.
% ref data/sources table?
However, the majority of the ploidy classifications were adopted from original studies with alternative sources of information (\eg geographic distribution, genus ploidy distribution) where ploidy was defined by authors that found evidence for it.
Since it is not clear whether diploidization can be detected under alternative ploidy classifications or even classifications based on chromosome number multiplicity at the genus level, we also fit the models without diploidization in order to test  whether the conclusions about diversification are sensitive to including diploidization.
As discussed by \citet{servedio_2014}, the presence or absence of a hypothesis can have an exploratory goal.
In our case the diploidization parameter (or its absence, $\delta=0$) in our models is an opportunity to explore an assumption that might be important but that is not the single definitive process to understand the interactions among polyploidy, breeding system, and diversification.

@@

Parameters for each of the 10 diversification models were performed using custom code in the RevBayes \citep{hoehna_2016} environment. %B: fixme: I think the intent here is  "Parameter estimation [...] was peformed"
Code for analyses and key results is available at \url{https://github.com/roszenil/solploidy}.
We included a correction for incomplete sampling in all analyses, based on assuming that the Solanaceae family has approximately 3,000 species ($s=595/3000$) as estimated by the Solanaceae Source project \citep{solsource}.
For all 10 models, we assumed that speciation and extinction parameters had log-normal prior distributions with means equal to the expected net diversification rate $(\text{number of taxa} / [2 \times \text{root age}])$ and standard deviation $0.5$.
Priors for parameters defining trait changes were assumed to be gamma distributed with parameters $k=0.5$ and $\theta=1$. 
For each model, an MCMC chain was run for 96 hours in the high-performance computational cluster at the Minnesota Supercomputing Institute, which allowed for 5,000 generations of burn-in and a minimum of 200,000 generations of MCMC for each of the 6 models. %B: the number of models changes here?
For each model, convergence and mixing of the MCMC was tested using the R library \texttt{coda} and the software package Tracer (see supplementary information for convergence plots). % fixme: eventually, add supp info fig ref

@@

We calculated the marginal likelihood for each of the 10 models in RevBayes \citep{hoehna_2016}.
Marginal likelihoods were calculated using 50 stepping stone steps under the methodology of \citet{xie_2010}.
Each stepping stone step was found by calculating 500 generations of burn-in followed by a total of 1,000 MCMC steps (\cref{table:marginallike}).
The calculation of each marginal likelihood ran for 24 hours on a high-performance computational cluster.

Using the marginal likelihood values, we calculated thirteen different Bayes factors. % E: it said "fourteen" but I only count thirteen? %B: fixme: my count too.
Six compared the models of polyploidy against each other (D/P and D/P--A/B, each with or without diploidization), one compared the breeding system models (I/C and I/C--A/B), and six compared the models with both traits (ID/CD/CP and ID/CD/CP--A/B, each with or without diploidization) (\cref{table:bayesfactors}).
%
Other comparisons between these models are not valid because the input data are different under the different state space codings (\cref{figure:stateclassifications}).
In essence, the D/P, I/C, and ID/CD/CP state spaces are not lumpable with respect to one another \citep{tarasov_2019}.
%R: Emma this is where the figure can be important to show that we have to make different input
% E: I think this short explanation is clear enough, especially with the figure.  I'm not sure whether the hierarchical concept is exactly applicable here.
%B: my TeX GUI (and my brain) underlined "lumpable", especially in the context of "with respect to one another." I think I get the gist, but I am not sure that I agree others will easily. I find it useful to imagine Spencer reading this, for example.

@@

\begin{table}
\begin{tabular}{@{}llcccccc@{}} \toprule
%\multicolumn{4}{r}{Models} \\ \cmidrule(r){3-5}
Model& Model& Ploidy & Diploidization & Breeding  & Hidden & Parameters & Marginal \\
& Type & &  &System & State & & Log- Likelihood \\
1. D/P &BiSSE &	Yes  &	Yes &	No	&No	& 6	& -1182.93 \\
2. D/P no $\delta$ & BiSSE  &	Yes & 	No	& No	& No & 	5 &	-1193.66\\
3. D/P- A/B  & HiSSE &	Yes &	Yes	&No &	Yes &	11 &	\textbf{-1145.69}\\
4. D/P no $\delta$-A/B &	HiSSE &	Yes & No &	No &	Yes &	10	&-1150.99\\
5. I/C &	BiSSE &	No &	 No	&Yes &	No &	 5 &  -1194.80 \\
6. I/C-A/B &	HiSSE &	No &	 No	&Yes &	Yes	& 10 & \textbf{-1155.37}\\
7. ID/P/CD & MuSSE &	Yes & 	Yes &	Yes &	No &	10 & -1344.50\\
8. ID/P/CD no $\delta$ &	MuSSE &	Yes & 	No &	Yes	&No &	9 &-1345.87\\
9. ID/P/CD-A/B & MuHiSSE &	Yes 	&Yes &	Yes &	Yes &	16 & \textbf{-1300.35} \\
10.ID/P/CD no $\delta$-A/B &	 MuHiSSE	 & Yes & 	No	&Yes &	Yes	&15 &-1303.55 \\ \bottomrule
\end{tabular}
\caption{Marginal likelihoods for the 10 diversification models proposed.}
\label{table:marginallike}
\end{table}
